---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# GlmSimulatoR
 
<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/GlmSimulatoR)](https://cran.r-project.org/package=GlmSimulatoR)
[![R build status](https://github.com/gmcmacran/GlmSimulatoR/workflows/R-CMD-check/badge.svg)](https://github.com/gmcmacran/GlmSimulatoR/actions)
[![Codecov test coverage](https://codecov.io/gh/gmcmacran/GlmSimulatoR/branch/master/graph/badge.svg)](https://codecov.io/gh/gmcmacran/GlmSimulatoR?branch=master)
<!-- badges: end -->

Often the first problem in understanding statistical models is finding good data. This package alleviates this by enabling you to create the data you need. With data in hand, you can answer questions. Is the estimated weight close to the true value? Does step wise search pick the correct variables? At what n does the sampling distribution normalize?

## What does ideal data for the linear model look like?

To answer this question, a simulated data set where Y follows a normal distribution is made.

```{r example1}
library(GlmSimulatoR)
library(ggplot2)

set.seed(1)
simdata <- simulate_gaussian(N = 100, weights = 2, xrange = 10, ancillary = 1) #GlmSimulatoR function
simdata %>% 
  print()
```

Then the data set is visualized.

```{r example1P2 }
ggplot(simdata, aes(x = X1, y = Y)) + 
  geom_point()
```

Voila The answer is found. See vignettes for more examples.
