---
title: "Count Data And Overdispersion"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Dealing With Right Skewed Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      results = "hide")
```

## Overview

For count response variables, the glm framework has two options. The poisson family and the negative binomial family. In this vignette, we will consider both and learn when to use one or the other.


## Getting Familiar With The Poisson Family

First lets get an idea of how the poisson family looks. If lambda is near 1, the poisson family is right skewed. As lambda increases, the poisson family becomes more gaussian. 

```{r Poisson Histogram, echo=TRUE}
library(stats)
library(MASS)
library(dplyr, warn.conflicts= FALSE)
library(ggplot2)
library(GlmSimulatoR)

set.seed(1)

#lambda = 1
poisson <- rpois(n = 10000, lambda = 1)
poissonDF <- as.data.frame(x=poisson)

ggplot(poissonDF, aes(x=poisson)) + 
  geom_histogram(bins = 100)

#lambda = 5
poisson <- rpois(n = 10000, lambda = 5)
poissonDF <- as.data.frame(x=poisson)

ggplot(poissonDF, aes(x=poisson)) + 
  geom_histogram(bins = 100)

#lambda = 10
poisson <- rpois(n = 10000, lambda = 10)
poissonDF <- as.data.frame(x=poisson)

ggplot(poissonDF, aes(x=poisson)) + 
  geom_histogram(bins = 100)

```

```{r echo=FALSE}
rm(poisson, poissonDF)

```


## GLM Poisson
The below histogram looks similar to the above histograms. Also, the Maximum Likelihood Estimation in the GLM framework was able to estimate the slopes well. 

```{r Poisson glm, echo=TRUE}

set.seed(1)
simdata <- simulate_poisson(N = 10000, weights = c(.5, 1))

#Response looks similar to above histograms
ggplot(simdata, aes(x=Y)) + 
  geom_histogram(bins = 100)

glmPoisson <- glm(Y ~ X1 + X2, data = simdata, family = poisson(link = "log"))
summary(glmPoisson)

```

```{r echo=FALSE}
rm(simdata, glmPoisson)
```

## Poisson Mean Variance Relationship
One important characteristic of the poisson family is $\mu$ equals $\sigma^2$. We see the sample statistics are nearly equal for different values of lambda.

```{r Poisson mu sigma, echo=TRUE}

set.seed(1)

#lambda = 1
poisson <- rpois(n = 10000, lambda = 1)
mean(poisson)
var(poisson)

#lambda = 5
poisson <- rpois(n = 10000, lambda = 5)
mean(poisson)
var(poisson)

#lambda = 10
poisson <- rpois(n = 10000, lambda = 10)
mean(poisson)
var(poisson)
```

```{r echo=FALSE}
rm(poisson)
```

## Poisson Vs Negative Binomial
For modeling, the main difference between poisson and the negative binomial is the extra parameter. The mean variance relationship is $\sigma^2$ equals $\mu + \mu^2/\theta$ for the negative binomial. Through $\theta$ many relationships are possible. 

When the response variable is a count, but $\mu$ does not equal $\sigma^2$, the poisson distribution is not applicable. Overdispersion can be detected by dividing the residual deviance by the degrees of freedom. If this quotient is much greater than one, the negative binomial distribution should be used. There is no hard cut off of "much larger than one", but a rule of thumb is 1.10 or greater is considered large.

## Getting Familiar With The Negative Binomial Family
When $\theta$ is large, $\mu + \mu^2/\theta$ is roughly equal to $\mu$ . Therefore the negative binomial will be similar to poisson distribution.

```{r Poisson Negative Binomial 1, echo=TRUE}
set.seed(1)

poisson <- rpois(n = 10000, lambda = 1)
poissonDF <- as.data.frame(x=poisson)

ggplot(poissonDF, aes(x=poisson)) + 
  geom_histogram(bins = 100)

negBin <- rnegbin(n = 10000, mu = 1, theta = 1000)
negBinDF <- as.data.frame(x=negBin)

ggplot(negBinDF, aes(x=negBin)) + 
  geom_histogram(bins = 100)

```

```{r echo=FALSE}
rm(poisson, poissonDF, negBin, negBinDF)
```


If $\theta$ is small, $\mu + \mu^2/\theta$ does not approximately equal $\mu$. The negative binomial will look different than poisson distribution. Note the difference in scale on the x axis and y axis.


```{r Poisson Negative Binomial 2, echo=TRUE}
set.seed(1)

poisson <- rpois(n = 10000, lambda = 1)
poissonDF <- as.data.frame(x=poisson)

ggplot(poissonDF, aes(x=poisson)) + 
  geom_histogram(bins = 100)

negBin <- rnegbin(n = 10000, mu = 1, theta = 1)
negBinDF <- as.data.frame(x=negBin)

ggplot(negBinDF, aes(x=negBin)) + 
  geom_histogram(bins = 100)
```

```{r echo=FALSE}
rm(poisson, poissonDF, negBin, negBinDF)
```

## GLM Negative Binomial
Lets create data where the flexibility of the negative binomial is needed and compare the poisson estimates to the negative binomial estimates. 

```{r Negative Binomial glm, echo=TRUE}

set.seed(1)
simdata <- simulate_negative_binomial(N = 10000, weights = c(.5, 1), ancillary = 5) #ancillary is theta.

#Response looks like a negative binomial distribution.
ggplot(simdata, aes(x=Y)) + 
  geom_histogram(bins = 200)

glmPoisson <- glm(Y ~ X1 + X2, data = simdata, family = poisson(link = "log"))
glmNB <- glm.nb(Y ~ X1 + X2, data = simdata, link = "log")


summary(glmPoisson)
summary(glmNB)

```

```{r echo=FALSE}
rm(simdata, glmPoisson, glmNB)
```
The estimated slopes are very similar. However, the standard errors are not. The fact $\mu$ does not equal $\sigma^2$ has caused a major under estimation of standard errors for the poisson GLM. For poisson models, it is important to look at residual deviance divided by the degrees of freedom. When this quotient is larger than 1, the negative binomial should be considered. In the above the quotient was 6.33 (63327 / 9997) for the poisson glm.
