---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


### Package Overview
Often the first problem in understanding statistical models is finding good data. This package alleviates this by enabling you to create the data you need. With data in hand, you can answer questions about models. Are the estimated weights close to the true values? Does step wise search pick the correct variables? At what n does the sampling distribution of weights normalize?

### Getting Familiar With The Package
All functions return a tibble. The only thing that changes between functions is the distribution of Y. In simulate gaussian, Y follows a gaussian distribution. In simulate_gamma, Y follows a gamma distribution.

### Example 1: Create Textbook Data for Linear Model

```{r example1}
library(GlmSimulatoR)
library(ggplot2)

set.seed(1)
simdata <- simulate_gaussian(N = 100, weights = 1, xrange = 10, ancillary = 1) #GlmSimulatoR function
ggplot(simdata, aes(x = X1, y = Y)) + 
  geom_point()

rm(simdata)
```

### Question 1: Is A Sample Size Of 200 Enough To Get Close Estimates Of The True Weights?
```{r setup}
set.seed(1)
simdata <- simulate_gaussian(N = 200, weights = c(1, 2, 3))

glmModel <- glm(Y ~ X1 + X2 + X3, data = simdata, family = gaussian(link = "identity"))
summary(glmModel)$coefficients

rm(glmModel)
```

In the above, we see the estimates are close to the weights argument. The mathematics behind the generalized linear model worked well.


### Question 2: Is The Model From Lm The Same As A Glm Model With Gaussian Family and Identity Link?
```{r}
library(GlmSimulatoR)
library(ggplot2)

set.seed(1)
simdata <- simulate_gaussian(link = "identity")

linearModel <- lm(Y ~ X1 + X2 + X3, data = simdata)
glmModel <- glm(Y ~ X1 + X2 + X3, data = simdata, family = gaussian(link = "identity"))

summary(linearModel)$coefficients
summary(glmModel)$coefficients

rm(simdata, linearModel, glmModel)
```

In the above, we see the coefficients and standard errors are the same between the linear model and the generalized linear model. This confirms the linear model is identical to a generalized linear model with gaussian family and identity link.
